Popularity Adjusted Block Models are Generalized Random Dot Product Graphs

### 2022-01-19 referee-response.Rmd

1. p4. What is a hollow graph?  
\textcolor{blue}{
We changed the terminology a bit here and throughout the paper 
and specify that the analysis assumes $G$ is an unweighted and undirected graph with no self-loops (no edge directly from itself to itself), 
so $A$ is binary and symmetric with zeros on the diagonal. 
}

2. p4. $G$ as a graph is non-oriented or undirected and $A$ is symmetric.  
\textcolor{blue}{
We have made minor edits clarifying the structure of $G$ and $A$. 
}

3. p4. Please define the set $[n]$.  
\textcolor{blue}{
We added $[n] = \{1, 2, ..., n\}$. 
}

4. p4. Definition 1. Do you consider a probability distribution on $z_1, ..., z_n$? Please give the range of variation of $(i, j)$ when you define $P_{ij}$.  
\textcolor{blue}{
The range of $(i, j)$ is now included in the definition. 
In regard to the probability distribution on the labels $z_1, ..., z_n$, none of our theorems use this, although some do require that each $n_k$ be sufficiently large, so it did not factor into our definitions and analyses. 
More specifically, the convergence rate of Theorem 5 depends on each $n_k$, and for convergence, we require that $n_k \to \infty$. 
While none of the theorems explicitly require this, in practice, we also require that each $n_k$ is large enough for the latent vectors in the $k^th$ community to span its subspace. 
}

### 2022-01-19 referee-response.tex

  \textcolor{blue}{
  The range of $(i, j)$ is now included in the definition. 
  In regard to the probability distribution on the labels $z_1, ..., z_n$, none of our theorems use this, although some do require that each $n_k$ be sufficiently large, so it did not factor into our definitions and analyses. 
  More specifically, the convergence rate of Theorem 5 depends on each $n_k$, and for convergence, we require that $n_k \to \infty$. 
  While none of the theorems explicitly require this, in practice, we also require that each $n_k$ is large enough for the latent vectors in the $k^th$ community to span its subspace. 

### 2022-01-26 referee-response.Rmd

12. p10, line 12. $c^{(i)}$ is not defined.  
\textcolor{blue}{
$c^{(i)}$ is the $i^{th}$ entry of column vector $c$. 
This has been added to the paper.
}

13. p10. I didn't understand the exposition of the SSC algorithm.  
\textcolor{blue}{
We have added more detail to our description of the SSC algorithm. 
In particular, we describe the intuition/motivation behind the algorithm. 
}  
\textcolor{blue}{The description of SSC now reads:
}  
\textcolor{blue}{
``The SSC algorithm can be described as follows: 
Given \(X \in \mathbb{R}^{n \times d}\) with vectors
\(x_i^\top \in \mathbb{R}^d\) as rows of \(X\), the optimization problem
\(c_i = \arg\min_{c} \|c\|_1\) subject to \(x_i = X^\top c\) and
\(c^{(i)} = 0\), where $c^{(i)}$ is the $i^{th}$ entry of $c$, is solved for each \(i \in [n]\). 
The solutions are collected into matrix
\(C = \bigl[ c_1 \mid \cdots \mid c_n \bigr]^\top\) to
construct an affinity matrix \(B = |C| + |C^\top|\). If each \(x_i\) lies
exactly on one of \(K\) subspaces, \(B\) describes an undirected graph
consisting of {\em at least} \(K\) disjoint subgraphs, i.e., \(B_{ij} = 0\) if \(x_i, x_j\) lie on different subspaces. 
The intuition here is that vectors that lie on the same subspace can be described as linear combinations of each other, assuming the number of vectors in the subspace is greater than the dimensionality of the subspace. 
Then once sparsity is enforced, for each $c_i$, its $j^{th}$ element $c_i^{(j)}$ is zero if $x_j$ belongs to a subspace that doesn't contain $x_i$, resulting in $B_{ij} = 0$. 
Thus, for each $c_i$, $c_i^{(j)}$ is zero if $x_i$ and $x_j$ belong to different subspaces and may be nonzero if they do. 
If \(X\) instead represents points near \(K\) subspaces with some noise, 
then this property may only hold approximately and 
a final graph partitioning step may be required 
(e.g., edge thresholding or spectral clustering).''
}

14. p11, line 31. "... SSC on the spectral embedding of $A$". Could you support this assertion by references?  
\textcolor{blue}{
There are no references for this, as it is a novel idea. 
SSC has been proposed for community detection for the PABM but for the rows of $A$ rather than the ASE of $A$ \citep{noroozi2019estimation}. 
The motivation for our implementation is from Theorem 1, which says that the PABM is a GRDPG for which the latent positions are comprised of subspaces. 
}  
\textcolor{blue}{
We have added, ``... SSC on the rows of the spectral embedding of $A$, since the embedding of $P$ consists of $K$ subspaces (Theorem 1), and so the embedding of $A$ will lie approximately on the $K$ subspaces \citep{rubindelanchy2017statistical}.''
}

### 2022-01-26 referee-response.tex

  p11, line 31. ``\ldots{} SSC on the spectral embedding of \(A\)''.
  Could you support this assertion by references?\\
  \textcolor{blue}{
  There are no references for this, as it is a novel idea. 
  SSC has been proposed for community detection for the PABM but for the rows of $A$ rather than the ASE of $A$ \citep{noroozi2019estimation}. 
  The motivation for our implementation is from Theorem 1, which says that the PABM is a GRDPG for which the latent positions are comprised of subspaces. 
  We have added, ``... SSC on the rows of the spectral embedding of $A$, since the embedding of $P$ consists of $K$ subspaces (Theorem 1), and so the embedding of $A$ will lie approximately on the $K$ subspaces \citep{rubindelanchy2017statistical}.''

### 2022-01-26 summary_minh_edited.tex
updated response to 14

However, because we do not observe $P$ but rather only the noisy
adjacency matrix $A \sim \BG(P)$, the natural approach then is
to perform SSC on the rows of the spectral embedding of \(A\), 
since they will still lie approximately on the $K$ subspaces 
\citep{rubindelanchy2017statistical}. 
We will show in Theorem~\ref{theorem5} that, with probability converging to
one as $n \rightarrow \infty$, the rows of the ASE of $A$ also
satisfy the subspace detection property. 
Theorem~\ref{theorem5} builds upon existing work by
\citet{rubindelanchy2017statistical} who describe the convergence
behavior of the ASE of \(A\) to that of \(\tilde{\Pi} XU\), and
\citet{jmlr-v28-wang13} who show the necessary conditions for the
subspace detection property to hold in noisy cases where the points lie
near subspaces. Finally we emphasize that while
\citet{noroozi2019estimation} also considered the use of SSC for
community recovery in PABM, they instead applied SSC to the rows of
\(A\) itself, foregoing
the embedding step altogether. It is however much harder to show that
the rows of $A$ satisfy the subspace detection property and thus, to
the best of our knowledge, there is currently no consistency result
regarding the application of SSC to the rows of $A$.

>>>>

However, because we do not observe $P$ but rather only the noisy
adjacency matrix $A \sim \BG(P)$, the natural approach then is
to perform SSC on the rows of the spectral embedding of $A$, 
since the embedding of $P$ consists of $K$ subspaces (Theorem~\ref{theorem2}), 
and so the embedding of $A$ will lie approximately on the $K$ subspaces 
\citep{rubindelanchy2017statistical}.
We will show in Theorem~\ref{theorem5} that, with probability converging to
one as $n \rightarrow \infty$, the rows of the ASE of $A$ also
satisfy the subspace detection property. 
Theorem~\ref{theorem5} builds upon existing work by
\citet{rubindelanchy2017statistical} who describe the convergence
behavior of the ASE of \(A\) to that of \(\tilde{\Pi} XU\), and
\citet{jmlr-v28-wang13} who show the necessary conditions for the
subspace detection property to hold in noisy cases where the points lie
near subspaces. Finally we emphasize that while
\citet{noroozi2019estimation} also considered the use of SSC for
community recovery in PABM, they instead applied SSC to the rows of
\(A\) itself, foregoing
the embedding step altogether. It is however much harder to show that
the rows of $A$ satisfy the subspace detection property and thus, to
the best of our knowledge, there is currently no consistency result
regarding the application of SSC to the rows of $A$.
